---
title: "All Numbers"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
source('/code/global.R')
source('/code/07_figures/CIPlot.R')
source('/code/05a_linearModel/LinearModelf.R')
load("/data/joinDataSources/cleanData/3.RData")
rawData <- repres$res
selOuts <- c( "FS_TotalGrayVol","FS_CortexVol")
```

This document computes all numbers for the day2day paper that are included in the main text.

##Participants
`r length(unique(rawData$subject))` subjects. Age is not included in the public data due to data security concerns.

Number of observations per particpant: `r min(table(rawData$subject))` to `r max(table(rawData$subject))`
```{r measuredura,echo=FALSE}
elapsed <- by(rawData$daysFromFirstScan,rawData$subject,max)

```
Time elapsed for participants ranged form  `r min(elapsed)/30` to `r max(elapsed)/30` month.

Total number of measurements: `r nrow(rawData)`

```{r measureintas,echo=FALSE}
projDir <- ''
timeDiffs <- rep(NA,nrow(rawData)) #actually less values but easier to undertand and code like this
for (i in 2:nrow(rawData)){
  if (rawData[i,'subject']==rawData[i-1,'subject'])
  timeDiffs[i] <- rawData[i,'scanDate']-rawData[i-1,'scanDate']
}
timeDiffs <- timeDiffs[!is.na(timeDiffs)]
```
Measurement Intervals: Minimum Difference: `r min(timeDiffs)`. Maximum Difference: `r max(timeDiffs)`, Mean Difference `r round(mean(timeDiffs),2)`, SD Difference `r round(sd(timeDiffs),2)`



#Missingness

##Brain
```{r missbrain,echo=FALSE}
sets <- getSettings(paper=1)
allMissBrain <- getMiss(rawData,sets$outcomes)
kable(allMissBrain$allmiss)
```

##Predictors
```{r misspreds,echo=FALSE}
#HELIUM IS WRONG BECAUSE IT IS IMPUTED BY ELISAS SCRIPT!!! MISSING HELIUM IS 7 or 7/281=2.49 %
sets <- getSettings(paper=1)
allMissCov <- getMiss(rawData,sets$preds)
kable(allMissCov$allmiss)
cat(sprintf('%d of %d covariates have missing',sum(allMissCov$allmiss$missRel>0),length(allMissCov$allmiss$missRel)))
#how much more missing than 5%
cat(sprintf('%d > 5 percent missing',sum(allMissCov$allmiss$missRel>5)))
df <- rawData[,c(sets$preds,'subject')]
dfOmit <- na.omit(df)
cat(sprintf('%d of %d rows dropped or %f by list wise deletion',nrow(dfOmit),nrow(df),nrow(dfOmit)/nrow(df)))
```



#ANOVA-like test

```{r anova,echo=FALSE}

load("/data/joinDataSources/cleanData/prepImpute/impute/repairFactors/linearModelImp/1.RData")
cRes <-repres$res
print(summary(cRes[[4]]))
print(summary(cRes[[5]]))
```

#Random Forest and LASSO
``` {r ml,echo=FALSE}
regPath <- file.path(projDir,'data/joinDataSources/cleanData/prepImpute/impute/repairFactors/linearModelImp/regularizedImp/paper1AllRegErrCor.RData')
rfPath <- file.path(projDir,'data/joinDataSources/cleanData/prepImpute/impute/repairFactors/addScanQuality/randomForestImp/1.RData')

```

#Within-Variance Improvements
``` {r within-improvements,echo=FALSE}
withinImprov <- function(result){
  within <- 1-result['error']
  improve <- result['errorRed']/within
  return(improve)
}

withinImprov2 <- function(errorBase,errorNew){
  within <- 1-errorBase
  withinNew <- 1-errorNew
  improve <- (within-withinNew)/within
  return(improve)
}

load("/data/joinDataSources/cleanData/prepImpute/impute/repairFactors/addScanQuality/getCorrelation/paper1.RData")
resultsBase <- repres$res$result
withinImprovDayGray <- withinImprov(resultsBase['FS_TotalGrayVol','daysFromFirstScan',]) 
withinImprovDayCort <- withinImprov(resultsBase['FS_CortexVol','daysFromFirstScan',])

load("/data/joinDataSources/cleanData/prepImpute/impute/repairFactors/addScanQuality/getCorrelation/paper1WDays.RData")
resultsDays <- repres$res$result
withinImprovTimeGray <- withinImprov(resultsDays['FS_TotalGrayVol','scanTime',]) 
withinImprovTimeCort <- withinImprov(resultsDays['FS_CortexVol','scanTime',])

file.path(projDir,'data/joinDataSources/cleanData/prepImpute/impute/repairFactors/linearModelImp/regularizedImp/paper1AllRegErrCor.RData')
load("/data/joinDataSources/cleanData/prepImpute/impute/repairFactors/addScanQuality/getCorrelation/paper1WTime.RData")
resultBoth <- repres$res$result
withinImprovTotalGray <- withinImprov2(resultsBase['FS_TotalGrayVol',1,'error'],resultBoth['FS_TotalGrayVol',1,'error'])
withinImprovTotalCort <- withinImprov2(resultsBase['FS_CortexVol',1,'error'],resultBoth['FS_CortexVol',1,'error'])

improvements <- data.frame(outcome=character(),variable=character(),improvement=numeric(),stringsAsFactors = FALSE)
improvements[1,] <- c("FS_TotalGrayVol","Day",round(100*withinImprovDayGray,digits=2))
improvements[2,] <- c("FS_CortexVol","Day",round(100*withinImprovDayCort,digits=2))
improvements[3,] <- c("FS_TotalGrayVol","Time",round(100*withinImprovTimeGray,digits=2))
improvements[4,] <- c("FS_CortexVol","Time",round(100*withinImprovTimeCort,digits=2))
improvements[5,] <- c("FS_TotalGrayVol","both",round(100*withinImprovTotalGray,digits=2))
improvements[6,] <- c("FS_CortexVol","both",round(100*withinImprovTotalCort,digits=2))
kable(improvements)
```


#Confidence Intervals

```{r CIs,echo=FALSE,results='asis'}
 sels <- c('lo95','est','hi95')
 CIs <- CIPlot(plot=FALSE)
 for (out in selOuts){
   for (type in c("raw","percent")){
     print(kable(CIs[[type]][[out]][,sels],digits = 2,caption=paste0(out, ' ' ,type)))
   }
 }
```

